\chapter*{Abstract}
\chaptermark{abstract}
\addcontentsline{toc}{chapter}{Abstract}

Many ecological communities show variation from place to place; understanding the causes of this variation is the goal of community ecology. Differences in community composition will be the result of both stochastic and deterministic processes. However, it is difficult to know to what degree, and under which circumstances, deterministic processes will shape community composition. In this thesis I combined observational and experimental approaches to quantify deterministic processes within a particular ecological community -- they phytotelmata of bromeliad plants. In my thesis I describe three studies at different scales of organization: 1) do organisms of different size respond equally to the same environmental gradient 2) how to predators interact to influence prey survival 3) what mechanisms underly the response of similar species to the same environmental gradient, bromeliad size. 

In Chapter 1, I tested an hypothesis developed from previous observational data - that smaller organisms respond less than larger ones to the same environmental gradient. I tested this hypothesis by selecting bromeliads from different habitats, homogenizing their bacteria, zooplankton and insect communities, then returning these bromeliads to their original habitats. I found that divergence in the bacterial communities was less than zooplankton and insects. In my second chapter, I examined ecological determinism on a smaller scale -- within a single trophic level (macroinvertebrate predators). Predation is a common source of ecological determinism, as it can both select for specific prey species and also change how these prey interact with the community (i.e.~by changing how the prey community cycles Nitrogen). In my final data chapter, I examine the way in which bromeliad macroinvertebrates respond to habitat area. This chapter has two parts -- first, I use a null model to demonstrate that species vary in their response to area more than can be explained by sampling effects alone. Then I use a detailed field experiment to demonstrate that for at least one such pair, a difference in abiotic tolerances may be the plausible mechanism. 

Together these results illustrate when, and to what degree, bromeliad communities respond to deterministic factors. All three chapters first demonstrate a pattern, testing it against a suitable null distribution, before attempting to quantify possible mechanisms with a field experiment. This combination of observation and experiment is an approach which can contribute to our understanding of how ecological systems work. 
%%% Local Variables:
%%% TeX-master: "thesis"
%%% TeX-PDF-mode: t
%%% End:
